---
output: github_document
---

# Rate calculation example

This is an R script of a reproducible example for mass aggregation of scientific data, as described in - 

> "Robust mass aggregation-based bedload flux prediction in flash flood conditions using an acoustic sensor" by Eran Halfi, Dror Paz, Kyle Stark, Ian Reid, Michael Dorman, Jonathan B. Laronne

Usage: replace the sample CSV file `bedload_mass_calibration_example.csv` file with a different one, following this structure. 

```
##                  date   weight level  Hp
## 1 2013-12-11 23:58:00 0.029397   0.0 0.0
## 2 2013-12-11 23:58:30 0.271035   0.2 0.1
## 3 2013-12-11 23:59:00 0.626709   0.4 0.1
## 4 2013-12-11 23:59:30 1.008832   0.4 0.1
## 5 2013-12-12 00:00:00 1.303752   0.3 0.1
## 6 2013-12-12 00:00:30 1.673890   0.4 0.1
```

## Reading data

```{r}
# Read file
dat = read.csv("bedload_mass_calibration_example.csv", skip = 1, stringsAsFactors = FALSE)
head(dat)

# Selecting relevant columns
dat = data.frame(
  date = dat[, 1],
  weight = dat[, 2],
  level = dat[, 3],
  Hp = dat[, 4]
)
head(dat)

# Filtering empty incomplete rows
dat = dat[complete.cases(dat), ]

# Formatting date-time
dat$date = as.POSIXlt(paste0(dat$date, ".00"), format = "%m/%d/%Y %H:%M:%S") 
```

## Rate calculation 

```{r}
# Set threshold
threshold = 4

# Consecutive differences
dat$change = c(0, diff(dat$weight))
head(dat)

# Cumulative differences, reset when threshold passed
dat$cumulative = NA
dat$cumulative[1] = 0
for(i in 2:nrow(dat)) {
  dat$cumulative[i] = ifelse(
    dat$cumulative[i-1] >= threshold, 
    dat$change[i], 
    dat$change[i] + dat$cumulative[i-1] 
    )
}
head(dat)

# Mark change points
dat$index = dat$cumulative > threshold
head(dat)

# Cumulative 'level' and 'Hp', with reset at change points
dat$level2 = NA
dat$Hp2 = NA
dat$Hp2_sum = NA
steps = which(dat$index)
for(i in 1:length(steps)) {
  if(i == 1) {w = 0:steps[i]} else {w = (steps[i-1]+1):steps[i]}  # Window to average
  dat$level2[steps[i]] = mean(dat$level[w])
  dat$Hp2[steps[i]] = mean(dat$Hp[w])
  dat$Hp2_sum[steps[i]] = sum(dat$Hp[w])
}
head(dat)

# Remove all rows which do not comprise change points
dat = dat[dat$index, c("date", "weight", "level2", "Hp2", "Hp2_sum")]
head(dat)

# Cumulative Hp2
dat$Hp2_cumulative = cumsum(dat$Hp2)
head(dat)

# Calculate flux
dat$change = c(NA, diff(dat$weight))
dat$difftime = c(NA, as.numeric(diff(as.POSIXlt(dat$date)), unit = "secs"))
dat$flux = (dat$change / 0.11) / dat$difftime
head(dat)

# Caculate Hp rate
dat$Hp2_rate = dat$Hp2 / dat$difftime
head(dat)
```

## Graphics

```{r}
# Plot 1
fit = lm(Hp2_cumulative ~ weight, dat)
s = summary(fit)
title = paste0("R2=", round(s$adj.r.squared, 2))
plot(Hp2_cumulative ~ weight, dat, main = title)
abline(fit, col = "red")

# Plot 2
fit = lm(flux ~ Hp2_rate, dat)
s = summary(fit)
title = paste0("R2=", round(s$adj.r.squared, 2))
plot(flux ~ Hp2_rate, dat, main = title)
abline(fit, col = "red")

# Plot 3
fit = lm(Hp2_rate ~ flux, dat)
s = summary(fit)
title = paste0("R2=", round(s$adj.r.squared, 2))
plot(Hp2_rate ~ flux, dat, main = title)
abline(fit, col = "red")

# Plot 4
plot(flux ~ as.numeric(date), dat, type = "l", ylim = range(c(dat$level2, dat$flux), na.rm=TRUE))
points(flux ~ as.numeric(date), dat)
lines(level2 ~ as.numeric(date), dat, col = "red")
points(level2 ~ as.numeric(date), dat, col = "red")
```

## Saving output

```{r}
write.csv(dat, "result.csv", row.names = FALSE)
```

















